HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series

### 2023-01-16 tmlr_submission/hermes_paper_submission/hermes_paper.tex
add tmlr submission

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\usepackage{amssymb,amsmath,amsthm}
\usepackage{xcolor}
\usepackage{pdfpages}
\modulolinenumbers[5]

\journal{Journal of \LaTeX\ Templates}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

\begin{frontmatter}

\title{HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author[mymainaddress,mysecondaryaddress]{\'Etienne DAVID}

%% or include affiliations in footnotes:
\author[mysecondaryaddress]{Jean BELLOT}

\author[mymainaddress]{Sylvain LE CORFF}

\address[mymainaddress]{\small  Samovar, T\'el\'ecom SudParis,  D\'epartement CITI, Institut Polytechnique de Paris, France.}
\address[mysecondaryaddress]{Heuritech, 71 Rue Réaumur, 75002 Paris, France.}

\begin{abstract}
Developing models and algorithms to draw causal inference for time series is a long standing statistical problem. It is crucial for many applications, in particular for fashion or retail industries, to make optimal inventory decisions and avoid massive wastes. By tracking thousands of fashion trends on social media with state-of-the-art computer vision approaches, we propose a new model for fashion time series forecasting. Our contribution is  twofold. We first provide publicly\footnote[1]{\url{http://files.heuritech.com/raw_files/f1_fashion_dataset.tar.xz}} an appealing fashion dataset gathering \numberts\ weekly fashion time series. As influence dynamics are the key of emerging trend detection, we associate with each time series an external weak signal representing behaviors of influencers. Secondly, to leverage such a complex and rich dataset, we propose a new hybrid forecasting model. Our approach combines per-time-series parametric models with seasonal components and a global recurrent neural network to include sporadic external signals. This hybrid model provides state-of-the-art results on the proposed fashion dataset, on the weekly time series of the M4 competition \cite{makridakis2018}, and illustrates the benefit of the contribution of external weak signals.
\end{abstract}

\begin{keyword}
Hybrid models, Recurrent neural networks, Time series.
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}
\label{sec:introduction}

Multivariate time series forecasting is a widespread statistical problem with  many applications, see for instance \cite{sarkka2013, douc2014, zucchini2017} and the numerous references therein.
Parametric generative models provide explainable predictions with statistical guarantees based on a precise modeling of the predictive distributions of new data based on a record of past observations.
Calibrating these models, for instance using maximum likelihood inference, often requires a fair amount of tuning to design a time series-specific model to provide  accurate forecasts and sharp confidence intervals.  Depending on the use case, statistical properties of the signal and the available data, many families of models have been proposed for time series.  The exponential smoothing model \cite{Brown1961}, the Trigonometric Box-Cox transform, ARMA errors, Trend, and Seasonal components model (TBATS) \cite{alysha2011}, or the ARIMA with the Box-Jenkins approach \cite{box2015} are for instance very popular parametric generative models.  Hidden Markov models (HMM) are also widespread and presuppose that available observations are defined using missing data describing the dynamical system. This hidden state is assumed to be a Markov chain such that at each time step the received observation is a random function of the corresponding latent data.  Although hidden states are modeled as a Markov chain, the observations arising therefrom have a complex statistical structure.
In various applications where signals exhibit non-stationarities such as trends and seasonality, classical HMM are not adapted. However, \cite{touron2017}  recently proposed seasonal HMM, assuming that transition probabilities between the states, as well as the emission distributions, are not constant in time but evolve in a periodic manner. Strong consistency results were established in \cite{touron2019} and Expectation Maximization based numerical experiments were proposed.
Although these works provide promising results, HMM are computationally expensive to train and are not yet well studied for seasonal  sequences with thousands of components.

In many fields, single or few time series have become thousands of sequences with various statistical properties. In this new context, classical time series specific statistical models show limitations when dealing with numerous heterogeneous data. Recurrent neural networks and recent sequence to sequence deep learning architectures offer very appealing numerical alternatives thanks to their capability of leveraging any kind of heterogeneous multivariate data, see for instance \cite{ hochreiter1997,vaswani2017, siami2018, li2019, lim2019,salinas2020}. The DeepAR model proposed in \cite{salinas2020} provides a global model from many time series based on a multi-layer recurrent neural network with LSTM cells. More recently, applications using the Transformer model have been proposed  \cite{li2019}. The Temporal Fusion Transformers (TFT) approach is a direct alternative to the DeepAR model \cite{lim2019}.  Unfortunately, all these solutions suffer from two main weaknesses. Firstly, many of them are black-boxes  as the final forecast usually does not come with a statistical guarantee  although a few recent works focused on measuring uncertainty in recurrent neural networks, see  \cite{martin2020}. Secondly, without a fine preprocessing and well chosen hyperparameters, these methods may lead to poor results and be outperformed by traditional statistical models, see \cite{makridakis2018}.

In this paper, we consider an emerging time series forecasting application referred to as {\em fashion trends prediction}. In fashion and retails industries, accurately anticipating consumers' needs is vital and wrong decisions can lead to massive wastes. Analysing social media pictures to detect what clothes people are wearing, it is possible to translate the visibility of fashion items over time into time series. Consequently, models and algorithms can be trained to accurately anticipate and predict consumers' behaviours. In \cite{ma2020}, a  dataset is provided using social media pictures and an image recognition framework to detect several clothes: 2000 fashion time series are proposed with a weekly seasonality. However, only 3 years of historical data is available (144 data points) that may not be sufficient for some statistical approaches. In \cite{ma2020}, another dataset is presented gathering 8000 fashion sequences with an historical available data  increased to 5 years. Nevertheless, only 120 values are available for each fashion time series and the overall volume remains low for a large part of the sequences resulting in a lot of noise and no clear patterns. In this paper, we propose a new fashion dataset overcoming the weaknesses of the two previous ones. Based on cutting-edge image recognition algorithms \cite{ren2015,chollet2017}, we built a large fashion dataset containing \numberts\ weekly sequences of fashion trends on social media with 5 years of historical data from 01-01-2015 to 30-12-2019. This dataset has very appealing properties:  all time series have the same length (261 data points), no missing value and no sparse time series even for niche trends. Concerning fashion dynamics, some of them appear to be really volatile with nonlinear changes of dynamics resulting from the emergence of new tendencies. In this context, understanding early signals of the apparition of a trend is one of the key to accurately forecast the future of the fashion. Consequently, the originality of our dataset comes from the fact that additional external weak signals are introduced. With our fashion expertise, we detected several groups of highly influential fashion users. Analyzing their specific behaviours on social media, we associate with each time series an external weak signal representing the same fashion trends on a sub-category of users. They are called weak signals because they are often alerts or events that are too sparse, or too incomplete to allow on their own an accurate estimation of their impact on the prediction of the target signal. Figure \ref{fig:oneemergingtrend} shows an example of fashion time series with its linked influencers weak signal. Exploring this new way of representing fashion, we aim at designing a model able to deal with such a large dataset, leveraging  complex external weak signals and finally providing the most accurate forecasts.

Recurrent neural networks are appealing to tackle our forecasting problem due to their capability of leveraging external data.  Recently, hybrid models combining deep neural network (DNN) architectures with widespread statistical models to deal with seasonality and trends have been proposed, see for instance  \cite{zhang2003,jianwei2019,bandara2020}. The approach providing the most striking results was proposed in  \cite{smyl2020} in the context of the M4 forecasting competition \cite{makridakis2020}.  Given a large dataset, a per-time-series multiplicative exponential smoothing model was introduced to estimate simple but fundamental components for each time series and compute a first prediction. Then a global recurrent neural network was trained on the entire dataset to correct errors of the previous exponential smoothing models.

@@

\subsection{Per-time-series predictors}
The time-series-specific predictors compute, for each sequence, a first $\lag$-ahead prediction based on the past. For all $1\leqslant n \leqslant N$, we note $\stat^n(.;\statparam^n)$ the $n$-th parametric model of the $n$-th sequence where $\statparam^n$ are  unknown parameters. Given the sequences $\{\ts^n_{1:T}\}_{1\leqslant n \leqslant N}$ and the estimated  parameters $\{\statparam^n\}_{1\leqslant n \leqslant N}$, the time-series-specific forecasts $\{\tspred^{pred,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
\begin{equation}
    \label{eq:predictors}
    \tspred^{pred,n}_{T+i|T} = \stat^n(\ts^n_{1:T};\statparam^n)_i\,.
\end{equation}
During the M4 competition, the hybrid model of \cite{smyl2020} was based on a multiplicative exponential smoothing model as the time-series-specific predictor. However, on sporadic time series, this choice leads to poor results and instability. In this paper, a more general framework able to deal with any kind of per-time-series models is provided. In Section~\ref{sec:exp}, two versions of our framework are proposed. The first one is based on an  exponential smoothing as a reference similar to the baseline \cite{smyl2020} and the second one uses a TBATS model \cite{alysha2011} which provides better results as this parametric model includes  Fourier representations with time varying coefficients, and ARMA error correction.

@@

\section{Fashion dataset with external weak signals}
\label{sec:dataset}

@@

Social media have appeared as an impressive data source to follow the evolution of fashion over the time. Looking at a specific trend, social media can provide where and when this trend had been worn at first and how it spread all over the world. To do it automatically for thousands of trends, the following methodology was introduced. In the first place, a complete image dataset of 150 millions pictures is collected from different social media such as Instagram or Weibo. We targeted 5 strategic markets for the retail industry using posts localisation: the United States, Europe, Japan, Brazil and China. 
The second step consists in creating a powerful visual recognition framework to be able to detect clothes details on pictures like the type of clothing, the form, the size, the color, the texture, etc. To do so, the following framework is designed.
\begin{enumerate}
    \item First, an object detection model is trained to detect the position, the size and the general type of possible multiple fashion items on a picture. This localization model is based on the Faster-RCNN architecture introduced in \cite{ren2015}. Starting from weights trained on MS-COCO \cite{lin2014}, a fine-tune of the model is done on our data with a standard setup following the original paper.
    \item  Additionally, several visual recognition models are trained at classifying a rich collection of 350 fashion details. We train one classifier for each category of fashion item: one for pants, another for tops, a third for shoes, etc. These models are all based on the Xception architecture introduced in \cite{chollet2017}. So as to trained them, large amount of social media pictures (between 200k and 800k training images depending on the category) have been manually tagged to constitute meaningful training datasets depending on the classification task. Architectures are first initialized with public weights trained on ImageNet \cite{russakovsky2014} and then fine tuned on the manually labeled dataset corresponding to their task.
\end{enumerate}

@@

Due to the increasing use of social media and continuous changes of users' behaviours, a normalization step is applied to the raw sequences in order to remove bias. Thus, we define the following normalizing signal $\Tilde{y}^{c,g,m}$. This signal represents the global sequence of the cloth type $c$ for the gender $g$ on market $m$ (e.g the evolution of the skirts in general for female in Europe). With the R package \texttt{stats}, the Seasonal-Trend decomposition using LOESS  \cite{cleveland1990} is used to remove the seasonal component of $\Tilde{y}^{c,g,m}$. The resulting deseasonalized signal is called  $\bar{y}^{c,g,m}$. Finally, for any fashion trend $i$, the following normalized sequence is defined for all $ 0 \leq t \leq T$:
\begin{equation}
    y^{i}_t = \frac{y^{c,g,m,i}_{t}}{\bar{y}^{c,g,m}_t}\,,
\end{equation}
where T denotes the number of available time steps. The time series $y^{c,g,m,i}$ is divided by the deseasonalized signal $\bar{y}^{c,g,m}$ and not $\Tilde{y}^{c,g,m}$ in order to avoid removing the seasonality of all the fashion trend sequences. With this normalizing step, a majority of the social media bias are removed and the final normalized sequences are expressed in share of category. As an illustration, an example of the normalization process is displayed in Figure~\ref{fig:normalization}. The raw Jersey Top trend for females in China is divided by the deseasonalized global Top trend for females in China.

@@

In theoretical fashion dynamics \cite{rogers1962}, different categories of adopters follow a trend in succession, resulting in several adoption waves.
So as too catch the early signal of the emergence of a trend, 6000 social media influencers were selectioned by hand by fashion experts. Aggregating them, a specific ``fashion-oriented`` panel is created. With the same methodology as for the main panel described in Section~\ref{sec:dataset:a} and Section~\ref{sec:dataset:b}, a normalized time series representing each fashion trend on this specific population is created. We named \textit{fashion-forwards} this weak signal.  For all fashion sequence $\{y^i_t\}_{1 \leq t \leq T}$, let $\{\ts^{f,i}_t\}_{1 \leq t \leq T}$ be the normalized sequence representing the behaviours of influencers regarding the fashion trend $i$. As we want to detect shifts between the main signal and the fashion forward signal, the following input is computed for the hybrid model: for all $t \in \{1,\ldots,T\}$ and any fashion trend $i$,
$$
\ws^{f,i}_{t} = \frac{\ts_t^{f,i}}{\ts_t^{f,i}+\ts_t^{i}}\,.
$$
where T denotes the number of available time steps. Values close to 0.5 indicate a similar behaviour between the influencers panel and the general panel. For instance, an impressive emerging fashion shoes trend with its \textit{fashion-forwards} weak signal is represented in Figure~\ref{fig:oneemergingtrend}.

@@

For the first parametric per-time-series models, existing Python  libraries named \texttt{statsmodels} and  \texttt{tbats} are used to estimate the different parameters $\statparam^n$.  Depending of the choice of local parametric models, two versions of HERMES are proposed. The first one uses as predictors an additive exponential smoothing model as a reference close to \cite{smyl2020}. The second one uses the TBATS model of \cite{alysha2011} and  achieves the highest accuracy results on the fashion dataset. The neural network architecture is composed of 3 LSTM layers of shape 50 and a final Dense layer to provide the correct output dimension. A classical Adam optimizer is used with a learning rate set at 0.001 or 0.005, the batch size is fixed to 64 and the loss function is defined as follows:
$$
\ell(\ts^n_{T+1:T+\lag},\tspred^n_{T+1:T+\lag|T}) = \frac{1}{\meants^n_T}\sum_{i=1}^{\lag}|\ts^n_{T+i} - \tspred^n_{T+i|T}|\,.
$$
This choice of $\mathrm{L}_1$ loss function is motivated by its robustness to outliers which accounts for some time series in the fashion industry with very specific behaviors. The loss and previous parameters are all set with a complete grid search. See \ref{sec:fashiongridsearch} for additional results concerning the loss function choice and \ref{sec:m4gridsearch} for a complete grid search example. The code is developed in Python using the Tensorflow library. It allows the use of GPU to speed up the training process.

@@

As benchmarks, several widespread statistical methods and deep learning approaches were selected. Using the R package \texttt{forecast} and the Python packages \texttt{statsmodels},  \texttt{tbats}, for each time series, predictions are computed with the following methods: \textit{snaive}, \textit{ets}, \textit{stlm}, \textit{thetam}, \textit{tbats} and \textit{auto.arima}. The forecast of the \textit{snaive} method is only the repetition of the last past period. The \textit{ets} model is an additive exponential smoothing with a level component and a seasonal component. The \textit{stlm} approach uses a multiplicative decomposition and models the seasonally adjusted time series with an exponential smoothing model. The \textit{Thetam} model decomposes the original signal in $\theta$-lines, predicts each one separately and recomposes them to produce the final forecast and \textit{tbats} uses a trigonometrical seasonality. Finally, \textit{auto.arima} is the R implementation of the ARIMA model with an automatic selection of the best parameters. A complete description and references for these models can be found in \cite{hyndman2020}. As a deep learning approach, a full LSTM (\textit{lstm}) neural network composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52 is considered.
Two versions of HERMES are proposed. They are called respectively \textit{hermes-ets} and \textit{hermes-tbats} according to the per-time-series model choice. Moreover, two versions with the inclusion of the weak signals (ws) are proposed. They are referred to as \textit{hermes-ets-ws} and \textit{hermes-tbats-ws}. In order to provide a fair comparison, a \textit{lstm} with the weak signals named \textit{lstm-ws} is trained.

@@

Considering the new HERMES approach, \textit{hermes-tbats} and \textit{hermes-tbats-ws} slightly outperform the alternatives in terms of MASE and are stable across the different trainings. Regarding \textit{hermes-ets},   although it is very similar to the baseline \cite{smyl2020}, its accuracy remains low in comparison to the \textit{lstm} benchmark or HERMES using TBATS.

@@

We also assessed the performance of HERMES using the M4 weekly dataset \cite{makridakis2020}. The M4 dataset gathers 359 weekly time series and has 3 main differences compared to the proposed fashion dataset. Firstly, sequences do not have the same length with sequences lying between 93 and 2610 time steps. Secondly, the 359 time series come from different sectors such that finance or Industry. Accordingly, they have very distinct scales and dynamics. Thirdly, compared to the previous fashion application, the time horizon of the prediction is set to 13 for the weekly dataset and no additional external signals are provided.

@@

\appendix

>>>>

\documentclass[10pt]{article} % For LaTeX2e
\usepackage{tmlr}
% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{tmlr}
% To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
%\usepackage[preprint]{tmlr}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.
\input{math_commands.tex}

\usepackage{hyperref}
\usepackage{url}
\usepackage{amssymb,amsmath,amsthm}
\usepackage[pdftex]{graphicx}

@@

\title{HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series}

% Authors must not appear in the submitted version. They should be hidden
% as long as the tmlr package is used without the [accepted] or [preprint] options.
% Non-anonymous submissions will be rejected without review.

\author{\name \'Etienne David \email etienne.david@heuritech.com \\
      \addr SAMOVAR, Télécom SudParis,\\
      Institut Polytechnique de Paris, 91120 Palaiseau, France
      \AND
      \name Jean Bellot \email jean.bellot@heuritech.com \\
      \addr Heuritech, \\
      6 Rue de Braque, 75003 Paris
      \AND
      \name Sylvain Le Corff \email sylvain.lecorff@gmail.com\\
      \addr LPSM, \\
      Sorbonne Université, UMR CNRS 8001, 75005, Paris
      }

% The \author macro works with any number of authors. Use \AND 
% to separate the names and addresses of multiple authors.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\def\month{01}  % Insert correct month for camera-ready version
\def\year{2023} % Insert correct year for camera-ready version
\def\openreview{\url{https://openreview.net/forum?id=XXXX}} % Insert correct link to OpenReview for camera-ready version

@@

\maketitle

\begin{abstract}
Developing models and algorithms to draw causal inference for time series is a long standing statistical problem. It is crucial for many applications, in particular for fashion or retail industries, to make optimal inventory decisions and avoid massive wastes. By tracking thousands of fashion trends on social media with state-of-the-art computer vision approaches, we propose a new model for fashion time series forecasting. Our contribution is  twofold. We first provide publicly\footnote[1]{\url{http://files.heuritech.com/raw_files/f1_fashion_dataset.tar.xz}} an appealing fashion dataset gathering \numberts\ weekly fashion time series. As influence dynamics are the key of emerging trend detection, we associate with each time series an external weak signal representing behaviors of influencers. Secondly, to leverage such a complex and rich dataset, we propose a new hybrid forecasting model. Our approach combines per-time-series parametric models with seasonal components and a global recurrent neural network to include sporadic external signals. This hybrid model provides state-of-the-art results on the proposed fashion dataset, on the weekly time series of the M4 competition \citep{makridakis2018}, and illustrates the benefit of the contribution of external weak signals.
\end{abstract}

\section{Introduction}
\label{sec:introduction}
Multivariate time series forecasting is a widespread statistical problem with  many applications, see for instance \citet{sarkka2013, douc2014, zucchini2017} and the numerous references therein.
Parametric generative models provide explainable predictions with statistical guarantees based on a precise modeling of the predictive distributions of new data based on a record of past observations.
Calibrating these models, for instance using maximum likelihood inference, often requires a fair amount of tuning to design a time series-specific model to provide  accurate forecasts and sharp confidence intervals.  Depending on the use case, statistical properties of the signal and the available data, many families of models have been proposed for time series.  The exponential smoothing model \citep{Brown1961}, the Trigonometric Box-Cox transform, ARMA errors, Trend, and Seasonal components model (TBATS) \citep{alysha2011}, or the ARIMA with the Box-Jenkins approach \citep{box2015} are for instance very popular parametric generative models.  Hidden Markov models (HMM) are also widespread and presuppose that available observations are defined using missing data describing the dynamical system. This hidden state is assumed to be a Markov chain such that at each time step the received observation is a random function of the corresponding latent data.  Although hidden states are modeled as a Markov chain, the observations arising therefrom have a complex statistical structure.
In various applications where signals exhibit non-stationarities such as trends and seasonality, classical HMM are not adapted. However, \citet{touron2017}  recently proposed seasonal HMM, assuming that transition probabilities between the states, as well as the emission distributions, are not constant in time but evolve in a periodic manner. Strong consistency results were established in \citet{touron2019} and Expectation Maximization based numerical experiments were proposed.
Although these works provide promising results, HMM are computationally expensive to train and are not yet well studied for seasonal  sequences with thousands of components.

In many fields, single or few time series have become thousands of sequences with various statistical properties. In this new context, classical time series specific statistical models show limitations when dealing with numerous heterogeneous data. Recurrent neural networks and recent sequence to sequence deep learning architectures offer very appealing numerical alternatives thanks to their capability of leveraging any kind of heterogeneous multivariate data, see for instance \citet{ hochreiter1997,vaswani2017, siami2018, li2019, lim2019,salinas2020}. The DeepAR model proposed in \citet{salinas2020} provides a global model from many time series based on a multi-layer recurrent neural network with LSTM cells. More recently, applications using the Transformer model have been proposed  \citep{li2019}. The Temporal Fusion Transformers (TFT) approach is a direct alternative to the DeepAR model \citep{lim2019}.  Unfortunately, all these solutions suffer from two main weaknesses. Firstly, many of them are black-boxes  as the final forecast usually does not come with a statistical guarantee  although a few recent works focused on measuring uncertainty in recurrent neural networks, see  \citet{martin2020}. Secondly, without a fine preprocessing and well chosen hyperparameters, these methods may lead to poor results and be outperformed by traditional statistical models, see \citet{makridakis2018}.

In this paper, we consider an emerging time series forecasting application referred to as {\em fashion trends prediction}. In fashion and retails industries, accurately anticipating consumers' needs is vital and wrong decisions can lead to massive wastes. Analysing social media pictures to detect what clothes people are wearing, it is possible to translate the visibility of fashion items over time into time series. Consequently, models and algorithms can be trained to accurately anticipate and predict consumers' behaviours. In \citet{ma2020}, a  dataset is provided using social media pictures and an image recognition framework to detect several clothes: 2000 fashion time series are proposed with a weekly seasonality. However, only 3 years of historical data is available (144 data points) that may not be sufficient for some statistical approaches. In \citet{ma2020}, another dataset is presented gathering 8000 fashion sequences with an historical available data  increased to 5 years. Nevertheless, only 120 values are available for each fashion time series and the overall volume remains low for a large part of the sequences resulting in a lot of noise and no clear patterns. In this paper, we propose a new fashion dataset overcoming the weaknesses of the two previous ones. Based on cutting-edge image recognition algorithms \citep{ren2015,chollet2017}, we built a large fashion dataset containing \numberts\ weekly sequences of fashion trends on social media with 5 years of historical data from 01-01-2015 to 30-12-2019. This dataset has very appealing properties:  all time series have the same length (261 data points), no missing value and no sparse time series even for niche trends. Concerning fashion dynamics, some of them appear to be really volatile with nonlinear changes of dynamics resulting from the emergence of new tendencies. In this context, understanding early signals of the apparition of a trend is one of the key to accurately forecast the future of the fashion. Consequently, the originality of our dataset comes from the fact that additional external weak signals are introduced. With our fashion expertise, we detected several groups of highly influential fashion users. Analyzing their specific behaviours on social media, we associate with each time series an external weak signal representing the same fashion trends on a sub-category of users. They are called weak signals because they are often alerts or events that are too sparse, or too incomplete to allow on their own an accurate estimation of their impact on the prediction of the target signal. Figure \ref{fig:oneemergingtrend} shows an example of fashion time series with its linked influencers weak signal. Exploring this new way of representing fashion, we aim at designing a model able to deal with such a large dataset, leveraging  complex external weak signals and finally providing the most accurate forecasts.

Recurrent neural networks are appealing to tackle our forecasting problem due to their capability of leveraging external data.  Recently, hybrid models combining deep neural network (DNN) architectures with widespread statistical models to deal with seasonality and trends have been proposed, see for instance  \citet{zhang2003,jianwei2019,bandara2020}. The approach providing the most striking results was proposed in  \citet{smyl2020} in the context of the M4 forecasting competition \citep{makridakis2020}.  Given a large dataset, a per-time-series multiplicative exponential smoothing model was introduced to estimate simple but fundamental components for each time series and compute a first prediction. Then a global recurrent neural network was trained on the entire dataset to correct errors of the previous exponential smoothing models.

@@

\subsection{Per-time-series predictors}
The time-series-specific predictors compute, for each sequence, a first $\lag$-ahead prediction based on the past. For all $1\leqslant n \leqslant N$, we note $\stat^n(.;\statparam^n)$ the $n$-th parametric model of the $n$-th sequence where $\statparam^n$ are  unknown parameters. Given the sequences $\{\ts^n_{1:T}\}_{1\leqslant n \leqslant N}$ and the estimated  parameters $\{\statparam^n\}_{1\leqslant n \leqslant N}$, the time-series-specific forecasts $\{\tspred^{pred,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
\begin{equation}
    \label{eq:predictors}
    \tspred^{pred,n}_{T+i|T} = \stat^n(\ts^n_{1:T};\statparam^n)_i\,.
\end{equation}
During the M4 competition, the hybrid model of \citet{smyl2020} was based on a multiplicative exponential smoothing model as the time-series-specific predictor. However, on sporadic time series, this choice leads to poor results and instability. In this paper, a more general framework able to deal with any kind of per-time-series models is provided. In Section~\ref{sec:exp}, two versions of our framework are proposed. The first one is based on an  exponential smoothing as a reference similar to the baseline \citet{smyl2020} and the second one uses a TBATS model \citep{alysha2011} which provides better results as this parametric model includes  Fourier representations with time varying coefficients, and ARMA error correction.

@@

section{Fashion dataset with external weak signals}
\label{sec:dataset}

@@

Social media have appeared as an impressive data source to follow the evolution of fashion over the time. Looking at a specific trend, social media can provide where and when this trend had been worn at first and how it spread all over the world. To do it automatically for thousands of trends, the following methodology was introduced. In the first place, a complete image dataset of 150 millions pictures is collected from different social media such as Instagram or Weibo. We targeted 5 strategic markets for the retail industry using posts localisation: the United States, Europe, Japan, Brazil and China. 
The second step consists in creating a powerful visual recognition framework to be able to detect clothes details on pictures like the type of clothing, the form, the size, the color, the texture, etc. To do so, the following framework is designed.
\begin{enumerate}
    \item First, an object detection model is trained to detect the position, the size and the general type of possible multiple fashion items on a picture. This localization model is based on the Faster-RCNN architecture introduced in \citet{ren2015}. Starting from weights trained on MS-COCO \citep{lin2014}, a fine-tune of the model is done on our data with a standard setup following the original paper.
    \item  Additionally, several visual recognition models are trained at classifying a rich collection of 350 fashion details. We train one classifier for each category of fashion item: one for pants, another for tops, a third for shoes, etc. These models are all based on the Xception architecture introduced in \citet{chollet2017}. So as to trained them, large amount of social media pictures (between 200k and 800k training images depending on the category) have been manually tagged to constitute meaningful training datasets depending on the classification task. Architectures are first initialized with public weights trained on ImageNet \citep{russakovsky2014} and then fine tuned on the manually labeled dataset corresponding to their task.
\end{enumerate}

@@

Due to the increasing use of social media and continuous changes of users' behaviours, a normalization step is applied to the raw sequences in order to remove bias. Thus, we define the following normalizing signal $\Tilde{y}^{c,g,m}$. This signal represents the global sequence of the cloth type $c$ for the gender $g$ on market $m$ (e.g the evolution of the skirts in general for female in Europe). With the R package \texttt{stats}, the Seasonal-Trend decomposition using LOESS  \citep{cleveland1990} is used to remove the seasonal component of $\Tilde{y}^{c,g,m}$. The resulting deseasonalized signal is called  $\bar{y}^{c,g,m}$. Finally, for any fashion trend $i$, the following normalized sequence is defined for all $ 0 \leq t \leq T$:
\begin{equation}
    y^{i}_t = \frac{y^{c,g,m,i}_{t}}{\bar{y}^{c,g,m}_t}\,,
\end{equation}
where T denotes the number of available time steps. The time series $y^{c,g,m,i}$ is divided by the deseasonalized signal $\bar{y}^{c,g,m}$ and not $\Tilde{y}^{c,g,m}$ in order to avoid removing the seasonality of all the fashion trend sequences. With this normalizing step, a majority of the social media bias are removed and the final normalized sequences are expressed in share of category. As an illustration, an example of the normalization process is displayed in Figure~\ref{fig:normalization}. The raw Jersey Top trend for females in China is divided by the deseasonalized global Top trend for females in China.

@@

In theoretical fashion dynamics \citep{rogers1962}, different categories of adopters follow a trend in succession, resulting in several adoption waves.
So as too catch the early signal of the emergence of a trend, 6000 social media influencers were selectioned by hand by fashion experts. Aggregating them, a specific ``fashion-oriented`` panel is created. With the same methodology as for the main panel described in Section~\ref{sec:dataset:a} and Section~\ref{sec:dataset:b}, a normalized time series representing each fashion trend on this specific population is created. We named \textit{fashion-forwards} this weak signal.  For all fashion sequence $\{y^i_t\}_{1 \leq t \leq T}$, let $\{\ts^{f,i}_t\}_{1 \leq t \leq T}$ be the normalized sequence representing the behaviours of influencers regarding the fashion trend $i$. As we want to detect shifts between the main signal and the fashion forward signal, the following input is computed for the hybrid model: for all $t \in \{1,\ldots,T\}$ and any fashion trend $i$,
$$
\ws^{f,i}_{t} = \frac{\ts_t^{f,i}}{\ts_t^{f,i}+\ts_t^{i}}\,.
$$
where T denotes the number of available time steps. Values close to 0.5 indicate a similar behaviour between the influencers panel and the general panel. For instance, an impressive emerging fashion shoes trend with its \textit{fashion-forwards} weak signal is represented in Figure~\ref{fig:oneemergingtrend}.

@@

For the first parametric per-time-series models, existing Python  libraries named \texttt{statsmodels} and  \texttt{tbats} are used to estimate the different parameters $\statparam^n$.  Depending of the choice of local parametric models, two versions of HERMES are proposed. The first one uses as predictors an additive exponential smoothing model as a reference close to \citet{smyl2020}. The second one uses the TBATS model of \citet{alysha2011} and  achieves the highest accuracy results on the fashion dataset. The neural network architecture is composed of 3 LSTM layers of shape 50 and a final Dense layer to provide the correct output dimension. A classical Adam optimizer is used with a learning rate set at 0.001 or 0.005, the batch size is fixed to 64 and the loss function is defined as follows:
$$
\ell(\ts^n_{T+1:T+\lag},\tspred^n_{T+1:T+\lag|T}) = \frac{1}{\meants^n_T}\sum_{i=1}^{\lag}|\ts^n_{T+i} - \tspred^n_{T+i|T}|\,.
$$
This choice of $\mathrm{L}_1$ loss function is motivated by its robustness to outliers which accounts for some time series in the fashion industry with very specific behaviors. The loss and previous parameters are all set with a complete grid search. See \ref{sec:fashiongridsearch} for additional results concerning the loss function choice and \ref{sec:m4gridsearch} for a complete grid search example. The code is developed in Python using the Tensorflow library. It allows the use of GPU to speed up the training process.

@@

As benchmarks, several widespread statistical methods and deep learning approaches were selected. Using the R package \texttt{forecast} and the Python packages \texttt{statsmodels},  \texttt{tbats}, for each time series, predictions are computed with the following methods: \textit{snaive}, \textit{ets}, \textit{stlm}, \textit{thetam}, \textit{tbats} and \textit{auto.arima}. The forecast of the \textit{snaive} method is only the repetition of the last past period. The \textit{ets} model is an additive exponential smoothing with a level component and a seasonal component. The \textit{stlm} approach uses a multiplicative decomposition and models the seasonally adjusted time series with an exponential smoothing model. The \textit{Thetam} model decomposes the original signal in $\theta$-lines, predicts each one separately and recomposes them to produce the final forecast and \textit{tbats} uses a trigonometrical seasonality. Finally, \textit{auto.arima} is the R implementation of the ARIMA model with an automatic selection of the best parameters. A complete description and references for these models can be found in \citet{hyndman2020}. As a deep learning approach, a full LSTM (\textit{lstm}) neural network composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52 is considered.
Two versions of HERMES are proposed. They are called respectively \textit{hermes-ets} and \textit{hermes-tbats} according to the per-time-series model choice. Moreover, two versions with the inclusion of the weak signals (ws) are proposed. They are referred to as \textit{hermes-ets-ws} and \textit{hermes-tbats-ws}. In order to provide a fair comparison, a \textit{lstm} with the weak signals named \textit{lstm-ws} is trained.

@@

Considering the new HERMES approach, \textit{hermes-tbats} and \textit{hermes-tbats-ws} slightly outperform the alternatives in terms of MASE and are stable across the different trainings. Regarding \textit{hermes-ets},   although it is very similar to the baseline \citet{smyl2020}, its accuracy remains low in comparison to the \textit{lstm} benchmark or HERMES using TBATS.

@@

We also assessed the performance of HERMES using the M4 weekly dataset \citep{makridakis2020}. The M4 dataset gathers 359 weekly time series and has 3 main differences compared to the proposed fashion dataset. Firstly, sequences do not have the same length with sequences lying between 93 and 2610 time steps. Secondly, the 359 time series come from different sectors such that finance or Industry. Accordingly, they have very distinct scales and dynamics. Thirdly, compared to the previous fashion application, the time horizon of the prediction is set to 13 for the weekly dataset and no additional external signals are provided.

@@

\bibliography{hermes_paper}
\bibliographystyle{tmlr}
